Paper

### 2024-12-14 paper/index.qmd
simpler ODE

# to change body font size:
# change this line in typst-template.typ:
# set text(font: "STIX Two Text", size: 24pt)
library(tidyverse)
library(abc)
source(file.path("..", "calibrate.R"))
# load(file.path("..", "out", "calibrate.RData"))
```

@@

-   $x_p(t) = 0$ if patient $p$ is not infected/colonised at time $t$,
-   $x_p(t) = 1$ if patient $p$ is infected/colonised at time $t$. It was assumed that infection/colonisation was not reversible during the simulation, i.e., that infected/colonised patients remained so until discharge. Hence, in contrast to a traditional susceptible-infected-recovered model [@anderson1991], the differential equations of this model were:

$$
\frac{dH_S}{dt} = A_S - Q\frac{H_S(t)}{H_I(t)} - DH_S(t)
$$

$$
\frac{dH_I}{dt} = A_I + Q\frac{H_S(t)}{H_I(t)} - DH_I(t)
$$

-   $dH_s/dt$ is the rate of
-   $H_S$ is the number of susceptible inpatients,
-   $H_I$ is the number of infected inpatients,
-   $A_S$ is the rate of new susceptible admissions,
-   $A_I$ is the rate of new infected admissions,
-   $Q$ is the transmission coefficient,
-   $D$ is the rate of patients that are discharged.

The hospital capacity was constantly kept at 100% occupancy (i.e., patients were immediately replaced on discharge), therefore:

$$
D = A_S + A_I
$$

At each time increment, the probability of a non-colonised/infected patient acquiring infection (i.e., transitioning from a non-infected/colonised to infected/colonised state) depended on: 1) a baseline risk of *de novo* infection/colonisation:

$$ 
P_\text{de novo}(t) = 1 - \exp {( - r_0 T(t) )}
$$

@@

@@

baselinerate <- seq(from=0.005, to=0.1, length.out = 100)
outbreakrate <- seq(0.005, 0.1, length.out = 100)
baselinecolonisation <- seq(0.9, 0.99, length.out = 100)
outbreakcolonisation <- seq(0.9, 0.99, length.out = 100)
rate_grid <- expand.grid(b_base = baselinerate,
                         b_out = outbreakrate,
                         c_base = baselinecolonisation,
                         c_out = outbreakcolonisation)
rate_grid <- rate_grid %>% 
  filter(b_base < b_out & c_base > c_out) %>% 
  slice_sample(n = 100)

@@

@@

plan(multisession)
ode_sims <- furrr::future_pmap_dfr(rate_grid,
        \(b_base, b_out, c_base, c_out) {
          N <- numeric(length(time))
          I <- numeric(length(time))
          N[1] <- initial_state["N"]
          I[1] <- initial_state["I"]

@@

            # Calculate derivatives
            dN <- p_t * parameters$A - parameters$A * prop_N - beta_t * N[t] * prop_I
            dI <- (1 - p_t) * parameters$A - parameters$A * prop_I + beta_t * N[t] * prop_I

@@

abc(c(target_baseline_rate, target_outbreak_rate), rate_grid, ode_sims[c("baseline", "outbreak")], tol=0.5, method='rejection') %>% 
  summary()

baselinerate <- 0.0236
outbreakrate <- 0.0482

# baselinecolonisation <- 1 - unique(single_outbreak_sim$`community-colonisation-rate`)
# outbreakcolonisation <- 1 - unique(single_outbreak_sim$`o-community-colonisation-rate`)
baselinecolonisation <- 0.9721
outbreakcolonisation <- 0.9403

@@

single_outbreak_one_seed %>% 
  pivot_longer(cols = c(`current-colonised`, `current_not_infected`), names_to = "compartment", values_to = "value") %>%
  group_by(compartment, `[step]`) %>% 
  summarise(m = mean(value)) %>% 
  ggplot(aes(x = `[step]`, y = m, color = compartment)) +
  geom_line() +
  labs(title = "Dynamics of Non-Infected and Infected Patients (Time-Varying Beta)",
       x = "Time",
       y = "Number of Patients",
       color = "Compartment") +
  geom_line(data = output_df, aes(x = time, y = N), color = "blue") + 
  geom_line(data = output_df, aes(x = time, y = I), color = "red")
```

>>>>

# to change body font size:
# change this line in typst-template.typ:
# set text(font: "STIX Two Text", size: 24pt)
library(tidyverse)
library(abc)
library(lhs)
source(file.path("..", "calibrate.R"))
# load(file.path("..", "out", "calibrate.RData"))
```

@@

-   $x_p(t) = 0$ if patient $p$ is not infected/colonised at time $t$,
-   $x_p(t) = 1$ if patient $p$ is infected/colonised at time $t$.

It was assumed that infection/colonisation was not reversible during the simulation, i.e., that infected/colonised patients remained so until discharge. At each time increment, the probability of a non-colonised/infected patient acquiring infection (i.e., transitioning from a non-infected/colonised to infected/colonised state) depended on: 1) a baseline risk of *de novo* infection/colonisation:

$$ 
P_\text{de novo}(t) = 1 - \exp {( - r_0 T(t) )},
$$

@@

A deterministic representation of the model was described in terms of the following differential equations, an adaptation of *SI* (susceptible--infection) models [@anderson1991;@keeling2011]:

$$
\frac{dS}{dt} = pA - D \cdot \frac{S}{S + I} - \beta S- rS
$$

$$
\frac{dI}{dt} = (1 - p)A - D \cdot \frac{I}{S + I} + \beta S + rS
$$

-   $dS/dt$ is the rate of change of susceptible/non-infected inpatients,
-   $S$ is the number of susceptible inpatients,
-   $I$ is the number of infected/colonised inpatients,
-   $A$ is the rate of new admissions,
-   $p$ is the proportion of new admissions that are susceptible/non-infected, while $(1 - p)$ is community colonisation/infection rate,
-   $\beta$ is the transmission coefficient,
-   $r$ is the rate of random (_de novo_) colonisation/infection,
-   $D$ is the rate of patient discharges.

The hospital capacity was constantly kept at 100% occupancy (i.e., patients were immediately replaced on discharge), therefore, $D = A$.

@@

randomLHS <- randomLHS(10000, 5)

baselinerate <- qunif(randomLHS[, 1], 0.0005, 0.01)
outbreakrate <- qunif(randomLHS[, 2], 0.0005, 0.01)
baselinecolonisation <- qunif(randomLHS[, 3], 0.9, 0.99)
outbreakcolonisation <- qunif(randomLHS[, 4], 0.9, 0.99)
r <- qunif(randomLHS[, 5], 0.0001, 0.001)
# rate_grid <- expand.grid(b_base = baselinerate,
#                          b_out = outbreakrate,
#                          c_base = baselinecolonisation,
#                          c_out = outbreakcolonisation,
#                          r = r)
rate_grid <- data.frame(b_base = baselinerate,
                        b_out = outbreakrate,
                        c_base = baselinecolonisation,
                        c_out = outbreakcolonisation,
                        r = r)
rate_grid <- rate_grid %>% 
  filter(b_base < b_out & c_base > c_out) %>% 
  slice_sample(n = 300)

@@

run_to <- max(single_outbreak_sim$`[step]`)

@@

plan(multisession)
ode_sims <- furrr::future_pmap_dfr(rate_grid,
        \(b_base, b_out, c_base, c_out, r) {
          N <- numeric(length(time))
          I <- numeric(length(time))
          N[1] <- initial_state["N"]
          I[1] <- initial_state["I"]

@@

            # Calculate derivatives
            dN <- p_t * parameters$A - parameters$A * prop_N - beta_t * N[t]  - r * N[t]
            dI <- (1 - p_t) * parameters$A - parameters$A * prop_I + beta_t * N[t]  + r * N[t]

@@

abc_single_outbreak <- abc(c(target_baseline_rate, target_outbreak_rate), rate_grid, ode_sims[c("baseline", "outbreak")], tol=0.025, method='rejection')

baselinerate <- mean(abc_single_outbreak$unadj.values[, "b_base"])
outbreakrate <- mean(abc_single_outbreak$unadj.values[, "b_out"])

# baselinecolonisation <- 1 - unique(single_outbreak_sim$`community-colonisation-rate`)
# outbreakcolonisation <- 1 - unique(single_outbreak_sim$`o-community-colonisation-rate`)
baselinecolonisation <- mean(abc_single_outbreak$unadj.values[, "c_base"])
outbreakcolonisation <- mean(abc_single_outbreak$unadj.values[, "c_out"])
r <- mean(abc_single_outbreak$unadj.values[, "r"])

@@

single_outbreak_one_seed %>% 
  pivot_longer(cols = c(`current-colonised`, `current_not_infected`), names_to = "compartment", values_to = "value") %>%
  filter(`[step]` > 31) %>%
  group_by(compartment, `[step]`) %>% 
  summarise(m = mean(value)) %>% 
  ggplot(aes(x = `[step]`, y = m, color = compartment)) +
  geom_line() +
  labs(title = "Dynamics of Non-Infected and Infected Patients (Time-Varying Beta)",
       x = "Time",
       y = "Number of Patients",
       color = "Compartment") +
  geom_line(data = output_df, aes(x = time, y = N), color = "blue") + 
  geom_line(data = output_df, aes(x = time, y = I), color = "red")
```
